---
title: "COVID VIZ"
author: "Jeffrey Brennan"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(warning = F)

library(ggplot2)
library(ggpubr)
library(ggrepel)
library(readr)
library(scales)
```


```{r}
confirmed = read.csv(url('https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv'))

deaths = read.csv(url('https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_US.csv'))
```

```{r}
confirmed$Type = 'confirmed'
deaths$Type = 'deaths'

texas_deaths = subset(deaths, Province_State == 'Texas')
texas_confirmed = subset(confirmed, Province_State == 'Texas')
texas_df = rbind(texas_confirmed, texas_deaths[, -12])
texas_df$Population = rep(texas_deaths$Population, times = 2)
# big_city_ids = texas_df[order(texas_df$Population, decreasing = T)[1:5], ][, 'UID']
# big_city_pops = texas_df[texas_df$UID %in% big_city_ids, 'Population']
# 
# # combine dfs
# big_city_df = rbind(confirmed[confirmed$UID %in% big_city_ids, ],
#                     deaths[deaths$UID %in% big_city_ids, -12])
# 
# big_city_df$Population = rep(big_city_pops, times = 2)
```



```{r}
# Convert to long format

# drop unused cols & fix names
texas_df = texas_df[, -c(2:4, 8, 11)]
colnames(texas_df)[3:6] = c('County', 'State', 'Lat', 'Long')

texas_long = reshape::melt(texas_df, id = c('UID', 'FIPS', 'County', 'State', 'Lat', 'Long', 'Type', 'Population'))
```


```{r}
# Clean long format data

colnames(texas_long)[9:10] = c('Date', 'Value')

# fix types
texas_long$Type = as.factor(texas_long$Type)
# texas_long$Location = droplevels(texas_long$Location)

# fix dates
texas_long$Date = gsub('X', '', texas_long$Date)
texas_long$Date = as.Date(as.character(texas_long$Date), format = '%m.%d.%y')
```


```{r}
# Convert days to normalized "Since 1st case.."
# sorted_df = big_city_long[order(big_city_long$Location), ] 
# 
# test = subset(sorted_df, Value != 0)
# 
# day_ones = test[!duplicated(test[c('Location', 'Type')]),]
# date_delay = as.integer(day_ones$Date - sorted_df$Date[1])
# 
# sorted_df$Day = (sorted_df$Date - sorted_df$Date[1]) + 1
# sorted_df$Day_norm = 0
# 
# 
# # TODO: try to figure out tidy solution
# for (i in seq(1:5)) {
#   sorted_df$Day_norm = 
#     with(sorted_df, 
#           ifelse(Location == levels(Location)[i] & Type == levels(Type)[1], Day - date_delay[2 * i - 1],
#                  ifelse(Location == levels(Location)[i] & Type == levels(Type)[2], Day - date_delay[2 * i], Day_norm)))
# }
# 
# 
# normed_df = subset(sorted_df, Day_norm > 0)
# ```
```

```{r}
# move deaths to own column (value -> cases, deaths)
texas_long$Deaths = subset(texas_long, Type == 'deaths')[, 'Value']
texas_long = subset(texas_long, Type == 'confirmed')
texas_long$Type = NULL
colnames(texas_long)[colnames(texas_long) == 'Value'] = 'Cases'
```


```{r}
# Add summary stats
texas_long$Cases_100k = (texas_long$Cases / texas_long$Population) * 100000
texas_long$Cases_Capita = (texas_long$Cases / texas_long$Population)^-1  # 1 in x...

texas_long$Deaths_100k = (texas_long$Deaths / texas_long$Population) * 100000
texas_long$Deaths_Capita =  (texas_long$Deaths / texas_long$Population)^-1  # 1 in x...
# replace inf with NA
# texas_long = do.call(data.frame, lapply(texas_long, function(x) replace(x, is.infinite(x), NA)))
```

```{r}
# save to csv
data_date = Sys.Date()
write.csv(texas_long, file = paste0('data/texas_df_', data_date, '.csv'), row.names = F)
```


<!-- ```{r} -->
<!-- # Split  -->


<!-- # Rename Locations for plotting -->
<!-- levels(normed_df$Location) = c('San Antonio', 'Dallas', 'Houston', 'Fort Worth', 'Austin') -->

<!-- normed_cases = subset(normed_df, Type == 'confirmed') -->
<!-- normed_deaths = subset(normed_df, Type == 'deaths') -->
<!-- ``` -->



<!-- ```{r} -->
<!-- # Relevel locations for plotting -->
<!-- newest_cases = subset(normed_cases, Date == max(Date)) -->
<!-- newest_deaths = subset(normed_deaths, Date == max(Date)) -->

<!-- case_location_order = as.character(newest_cases[order(newest_cases$Value, decreasing = T), 'Location']) -->
<!-- death_location_order = as.character(newest_deaths[order(newest_deaths$Value, decreasing = T), 'Location']) -->

<!-- normed_cases$Location = factor(normed_cases$Location, levels = death_location_order) -->
<!-- normed_deaths$Location = factor(normed_deaths$Location, levels = death_location_order) -->
<!-- ``` -->


# Cases

## Daily cases

### Log transformed

```{r, warning = FALSE}
ggplot(normed_cases,
       aes(x = Day_norm, y = Value, color = Location, group = Location)) +
  geom_point(size = 2) +
  geom_line(size = 1.5) +
  geom_text_repel(aes(label = ifelse(Day == max(Day), as.character(Location), NA)), size = 3.5,
                   nudge_x = 10, fontface = 'bold', min.segment.length = 0, segment.size=0.5, label.padding = 2) +
  scale_y_log10() +
  expand_limits(x = (max(normed_cases$Day_norm) * 1.1)) +
  labs(title = 'Daily Cases of COVID-19 in 5 Major Texas Cities',
       subtitle = '',
       y = 'log(Daily Cases)',
       x = '# Days since First Recorded Case') + 
  theme_pubr() + 
  theme(legend.position = 'none')
```

### Untransformed

```{r, warning = FALSE}
ggplot(normed_cases,
       aes(x = Day_norm, y = Value, color = Location, group = Location)) +
  geom_point(size = 2) +
  geom_line(size = 1.5) +
  geom_text_repel(aes(label = ifelse(Day == max(Day), paste0(as.character(Location), '\n', Value), NA)), size = 3.5,
                   nudge_x = 6, fontface = 'bold', min.segment.length = 0, segment.size=0.5, label.padding = 2) +
  expand_limits(x = (max(normed_cases$Day_norm) * 1.1)) +
  labs(title = 'Daily Cases of COVID-19 in 5 Major Texas Cities',
       subtitle = '',
       y = 'Daily Cases',
       x = '# Days since First Recorded Case') + 
  theme_pubr() + 
  theme(legend.position = 'none')
```


## Last 14 Days

```{r}
ggplot(subset(normed_cases, Date > (Sys.Date() - 14)),
       aes(x = Day_norm, y = Value, color = Location, group = Location)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  geom_text_repel(aes(label = ifelse(Day == max(Day), paste0(Location, '\n', Value), NA)), size = 3.5,
                 nudge_x = 1, fontface = 'bold', min.segment.length = 5, segment.size=0.5, label.padding = 2) +
  expand_limits(x = (max(normed_cases$Day_norm * 1.02))) +
  labs(title = 'Daily Cases of COVID-19 over Last 2 Weeks',
       subtitle = Sys.time(),
       y = 'Daily Cases',
       x = '# Days since First Recorded Case') + 
  theme_pubr() + 
  theme(legend.position = 'none')

ggsave('viz/case_summary.png', width = 8, height = 6, dpi = 600)
```


## Cumulative cases

## Log transformed

```{r, warning = FALSE}
ggplot(normed_cases %>% group_by(Location) %>% mutate(cv = cumsum(Value)),
       aes(x = Day_norm, y = cv, color = Location, group = Location)) +
  geom_point(size = 2) +
  geom_line(size = 1.5) +
  scale_y_log10() +
  geom_text_repel(aes(label = ifelse(Day == max(Day), as.character(Location), NA)), size = 3.5,
                   nudge_x = 6, fontface = 'bold', min.segment.length = 0, segment.size=0.5, label.padding = 2) +
  expand_limits(x = (max(normed_cases$Day_norm) * 1.1)) +
  labs(title = 'Cumulative Cases of COVID-19 in 5 Major Texas Cities',
       subtitle = '',
       y = 'Cumulative Cases',
       x = '# Days since First Recorded Case') + 
  theme_pubr() + 
  theme(legend.position = 'none')
```


## Untransformed

```{r, warning = FALSE}
ggplot(normed_cases %>% group_by(Location) %>% mutate(cv = cumsum(Value)),
       aes(x = Day_norm, y = cv, color = Location, group = Location)) +
  geom_point(size = 2) +
  geom_line(size = 1.5) +
  geom_text_repel(aes(label = ifelse(Day == max(Day), paste0(as.character(Location), '\n', cv), NA)), size = 3.5,
                   nudge_x = 6, fontface = 'bold', min.segment.length = 0, segment.size=0.5, label.padding = 2) +
  expand_limits(x = (max(normed_cases$Day_norm) * 1.1)) +
  labs(title = 'Cumulative Cases of COVID-19 in 5 Major Texas Cities',
       subtitle = '',
       y = 'Cumulative Cases',
       x = '# Days since First Recorded Case') + 
  theme_pubr() + 
  theme(legend.position = 'none')
```

# Deaths


## Daily Deaths

### Log transformed

```{r, warning = FALSE}
ggplot(normed_deaths,
       aes(x = Day_norm, y = Value, color = Location, group = Location)) +
  geom_point(size = 2) +
  geom_line(size = 1.5) +
  geom_text_repel(aes(label = ifelse(Day == max(Day), as.character(Location), NA)), size = 3.5,
                   nudge_x = 10, fontface = 'bold', min.segment.length = 0, segment.size=0.5, label.padding = 2) +
  scale_y_log10() +
  expand_limits(x = (max(normed_deaths$Day_norm) * 1.1)) +
  labs(title = 'Daily Deaths from COVID-19 in 5 Major Texas Cities',
       subtitle = '',
       y = 'log(Daily Deaths)',
       x = '# Days since First Recorded Death') + 
  theme_pubr() + 
  theme(legend.position = 'none')
```

### Untransformed

```{r, warning = FALSE}
ggplot(normed_deaths,
       aes(x = Day_norm, y = Value, color = Location, group = Location)) +
  geom_point(size = 2) +
  geom_line(size = 1.5) +
  geom_text_repel(aes(label = ifelse(Day == max(Day), paste0(as.character(Location), '\n', Value), NA)),
                  size = 3.5, nudge_x = 2, nudge_y = 10, fontface = 'bold', min.segment.length = 0,
                  segment.size=0.5, label.padding = 2) +
  expand_limits(x = (max(normed_deaths$Day_norm) * 1.1)) +
  labs(title = 'Daily Deaths from COVID-19 in 5 Major Texas Cities',
       subtitle = '',
       y = 'Daily Deaths',
       x = '# Days since First Recorded Death') + 
  theme_pubr() + 
  theme(legend.position = 'none')
```


## Last 14 Days

```{r}
ggplot(subset(normed_deaths, Date > (Sys.Date() - 14)),
       aes(x = Day_norm, y = Value, color = Location, group = Location)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  geom_text_repel(aes(label = ifelse(Day == max(Day), paste0(Location, '\n', Value), NA)), size = 3.5,
                 nudge_x = 0.5, nudge_y = 5, fontface = 'bold', min.segment.length = 5, segment.size=0,
                 label.padding = 0) +
  expand_limits(x = (max(normed_deaths$Day_norm * 1.02))) +
  labs(title = 'Daily Deaths from COVID-19 over Last 2 Weeks',
       subtitle = Sys.time(),
       y = 'Daily Deaths',
       x = '# Days since First Recorded Death') + 
  theme_pubr() + 
  theme(legend.position = 'none')

ggsave('viz/death_summary.png', width = 8, height = 6, dpi = 600)
```


## Cumulative Deaths

## Log transformed

```{r, warning = FALSE}
ggplot(normed_deaths %>% group_by(Location) %>% mutate(cv = cumsum(Value)),
       aes(x = Day_norm, y = cv, color = Location, group = Location)) +
  geom_point(size = 2) +
  geom_line(size = 1.5) +
  scale_y_log10() +
  expand_limits(x = (max(normed_deaths$Day_norm) * 1.1)) +
  labs(title = 'Cumulative Deaths from COVID-19 in 5 Major Texas Cities',
       subtitle = '',
       y = 'Cumulative Deaths',
       x = '# Days since First Recorded Death') + 
  theme_pubr() + 
  theme(legend.position = 'right')
```


## Untransformed

```{r, warning = FALSE}
ggplot(normed_deaths %>% group_by(Location) %>% mutate(cv = cumsum(Value)),
       aes(x = Day_norm, y = cv, color = Location, group = Location)) +
  geom_point(size = 2) +
  geom_line(size = 1.5) +
  geom_text_repel(aes(label = ifelse(Day == max(Day), paste0(as.character(Location), '\n', cv), NA)), size = 3.5,
                   nudge_x = 6, fontface = 'bold', min.segment.length = 0, segment.size=0.5, label.padding = 2) +
  expand_limits(x = (max(normed_deaths$Day_norm) * 1.1)) +
  labs(title = 'Cumulative Deaths from COVID-19 in 5 Major Texas Cities',
       subtitle = '',
       y = 'Cumulative Deaths',
       x = '# Days since First Recorded Death') + 
  theme_pubr() + 
  theme(legend.position = 'none')
```
